---
title: "CoP SAP 6.2"
author: "Allyson Mateja"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    includes:
      in_header: reformat_paragraph.tex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)
library(boot)
library(dplyr)
library(readxl)
library(GGally)
library(cowplot)
library(stringr)
library(ggpubr)
```



```{r, echo = FALSE}
#Functions 

#Function to get bootstrap measures of Spearman's correlation between 2 measures
boot.spearman_ci <- function(df, i) {
  df2 <- df[i, ]
  return(cor(df2$logx, df2$logy, use = "complete.obs"))
}

#Function to get Spearman's correlation and 95% CI by bootstrap
cor_return_boot <- function(logx, logy, R=5000) {
  #Get Spearman's correlation
  df1 <- data.frame(logx=logx, logy=logy)
  df1 <- df1[complete.cases(df1), ]
  df <- data.frame(cbind(rank(df1$logx), rank(df1$logy)))
  names(df) <- c("logx", "logy")
  cor_num <- cor.test(df$logx, df$logy)$estimate
  #Do 5000 bootstrap samples to calculate Spearman's CI, and get 2.5 %ile and 97.5 %ile as
  #95% bootstrap confidence bands
  boot_samp <- boot(df, boot.spearman_ci, R=R)$t
  lower <- quantile(boot_samp, 0.025)
  upper <- quantile(boot_samp, 0.975)
  
  cor_return <- data.frame(result = paste0(round(cor_num, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  
  return(cor_return)
}

#Function to get Spearman's correlation and 95% CI
cor_return <- function(logx, logy) {
  #Get Spearman's correlation
  #Determine ranks for X and Y
  
  df1 <- data.frame(logx=logx, logy=logy)
  df1 <- df1[complete.cases(df1), ]
  df <- data.frame(cbind(rank(df1$logx), rank(df1$logy)))
  names(df) <- c("logx", "logy")
  #Get upper and lower confidence bounds using Pearson's correlation on the ranks
  out <- cor.test(df$logx, df$logy, use="complete.obs")
  lower <- out$conf.int[1]
  upper <- out$conf.int[2]
  cor_num <- out$estimate
  #Return Spearman's cor and lower & upper CI bounds as data frame
  cor_return <- data.frame(result = paste0(round(cor_num, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  return(cor_return)
}

#This is a function for the pairs plot to remove the middle blank diagonal
gpairs_lower <- function(g){
  g$plots <- g$plots[-(1:g$nrow)]
  g$yAxisLabels <- g$yAxisLabels[-1]
  g$nrow <- g$nrow -1

  g$plots <- g$plots[-(seq(g$ncol, length(g$plots), by = g$ncol))]
  g$xAxisLabels <- g$xAxisLabels[-g$ncol]
  g$ncol <- g$ncol - 1

  g
}


#Function to make paired plots of immune markers
#Input dataframe and columns of immune markers on regular scale

immune_markers_pairs <- function(dat, cols, caption.num) {
  
  #For each manufacturer, plot a paired scatter plot for the immune correlates, with different colors for dose groups, log transform X and Y axes
  
  plots <- (gpairs_lower(ggpairs(dat, columns = cols, ggplot2::aes(col = Subgroup), upper  = list(continuous = "blank"), diag  = list(continuous = "blankDiag"), lower=list(continuous = wrap("points",size=0.5)), columnLabels = gsub('\\.|-|_', ' ', colnames(dat)[cols]), labeller = label_wrap_gen(18), legend = c(2, 1)) + scale_color_manual(values = c("red", "blue", "forestgreen", "tan1", "orchid1", "black", "purple3"), name = "Dose Group", drop=FALSE)  + theme_bw() + theme(strip.background=element_rect(fill="white"), strip.text = element_text(size=5), axis.text = element_text(size=5), axis.text.x=element_text(angle=90, vjust = 0.5, hjust=1), legend.position = "bottom") + ggtitle(dat[1, ]$manuf)))  
  
  plots[1,1]<-plots[1,1]+scale_y_continuous(trans="log10",limits=c(10,70000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[2,1]<-plots[2,1]+scale_y_continuous(trans="log10",limits=c(100,4000000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[3,1]<-plots[3,1]+scale_y_continuous(trans="log10",limits=c(100,800000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[4,1]<-plots[4,1]+scale_y_continuous(trans="log10",limits=c(5,90000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[5,1]<-plots[5,1]+scale_y_continuous(trans="log10",limits=c(5,70000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[6,1]<-plots[6,1]+scale_y_continuous(trans="log10",limits=c(5,35000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[7,1]<-plots[7,1]+scale_y_continuous(trans="log10",limits=c(5,15000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  
  
  plots[2,2]<-plots[2,2]+scale_y_continuous(trans="log10",limits=c(100,4000000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[3,2]<-plots[3,2]+scale_y_continuous(trans="log10",limits=c(100,800000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[4,2]<-plots[4,2]+scale_y_continuous(trans="log10",limits=c(5,90000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[5,2]<-plots[5,2]+scale_y_continuous(trans="log10",limits=c(5,70000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[6,2]<-plots[6,2]+scale_y_continuous(trans="log10",limits=c(5,35000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  plots[7,2]<-plots[7,2]+scale_y_continuous(trans="log10",limits=c(5,15000))+scale_x_continuous(trans="log10",limits=c(10,70000))
  
   plots[3,3]<-plots[3,3]+scale_y_continuous(trans="log10",limits=c(100,800000))+scale_x_continuous(trans="log10",limits=c(100,4000000))
  plots[4,3]<-plots[4,3]+scale_y_continuous(trans="log10",limits=c(5,90000))+scale_x_continuous(trans="log10",limits=c(100,4000000))
  plots[5,3]<-plots[5,3]+scale_y_continuous(trans="log10",limits=c(5,70000))+scale_x_continuous(trans="log10",limits=c(100,4000000))
  plots[6,3]<-plots[6,3]+scale_y_continuous(trans="log10",limits=c(5,35000))+scale_x_continuous(trans="log10",limits=c(100,4000000))
  plots[7,3]<-plots[7,3]+scale_y_continuous(trans="log10",limits=c(5,15000))+scale_x_continuous(trans="log10",limits=c(100,4000000))
  
  plots[4,4]<-plots[4,4]+scale_y_continuous(trans="log10",limits=c(5,90000))+scale_x_continuous(trans="log10",limits=c(100,800000))
  plots[5,4]<-plots[5,4]+scale_y_continuous(trans="log10",limits=c(5,70000))+scale_x_continuous(trans="log10",limits=c(100,800000))
  plots[6,4]<-plots[6,4]+scale_y_continuous(trans="log10",limits=c(5,35000))+scale_x_continuous(trans="log10",limits=c(100,800000))
  plots[7,4]<-plots[7,4]+scale_y_continuous(trans="log10",limits=c(5,15000))+scale_x_continuous(trans="log10",limits=c(100,800000))
  
  
  plots[5,5]<-plots[5,5]+scale_y_continuous(trans="log10",limits=c(5,70000))+scale_x_continuous(trans="log10",limits=c(5,90000))
  plots[6,5]<-plots[6,5]+scale_y_continuous(trans="log10",limits=c(5,35000))+scale_x_continuous(trans="log10",limits=c(5,90000))
  plots[7,5]<-plots[7,5]+scale_y_continuous(trans="log10",limits=c(5,15000))+scale_x_continuous(trans="log10",limits=c(5,90000))
  
  plots[6,6]<-plots[6,6]+scale_y_continuous(trans="log10",limits=c(5,35000))+scale_x_continuous(trans="log10",limits=c(5,70000))
  plots[7,6]<-plots[7,6]+scale_y_continuous(trans="log10",limits=c(5,15000))+scale_x_continuous(trans="log10",limits=c(5,70000))
  
  plots[7,7]<-plots[7,7]+scale_y_continuous(trans="log10",limits=c(5,15000))+scale_x_continuous(trans="log10",limits=c(5,35000))
  
  plots <- ggmatrix_gtable(plots) 
  
  p <- ggarrange(plots)
  annotate_figure(p, fig.lab = paste0("Figure 6.2.", caption.num), fig.lab.pos = "bottom.right")
}

#Function to make paired plots of immune markers with outcomes
#Input dataframe,  columns of immune markers on regular scale, and columns of outcomes on regular scale
markers_outcomes_pairs <- function(dat, Xcols, Ycols, caption.num) {
  #For each manufacturer, plot a paired scatter plot for the immune correlates x outcomes, with different colors for dose groups, log transform X and Y axes

  p <- gpairs_lower(ggpairs(dat, columns = c(Xcols, Ycols), ggplot2::aes(col = Subgroup), upper  = list(continuous = "blank"), diag  = list(continuous = "blankDiag"), lower=list(continuous = wrap("points",size=0.5)), columnLabels = gsub('\\.|-|_', ' ', colnames(dat)[c(Xcols, Ycols)]), labeller = label_wrap_gen(8), legend = c(2, 1)) + theme_bw() + theme(strip.background=element_rect(fill="white"), strip.text = element_text(size=5), axis.text = element_text(size=5), axis.text.x=element_text(angle=90, vjust = 0.5, hjust=1), legend.position = "bottom") + ggtitle(dat[1, ]$manuf) + scale_color_manual(values = c("red", "blue", "forestgreen", "tan1", "orchid1", "black", "purple3"), name = "Dose Group", drop=FALSE))
  
    
    #p produces a plot with all pairs of immune correlates and outcomes - however, we are not interested in pairs of immune correlates together or outcomes only
    #Get only pairs of immune correlates with outcomes
  p_y <- vector(mode = "list", length = length(Ycols))
  for (k in 1:length(Ycols)) {
    p_y[[k]] <- lapply(1:length(Xcols), function(j) getPlot(p, i = length(Xcols) + k-1, j = j))
  }
  
  new <- vector(mode = "list", length = length(Ycols)*length(Xcols))
  for (i in 1:length(Ycols)) {
    new[c(c(1:8) + (8*(i-1)))] <- p_y[[i]][c(1:8)]
  }
  
  xmin <- rep(c(10,10,100,100,5,5,5,5), 12)
  xmax <- rep(c(70000,70000,4000000,800000,90000,70000,35000,15000),12)
  ymin <- rep(c(200,100,200,100,200,100,15,15,25,25,25,25),each=8)
  ymax <- rep(c(4e7,5e8,2500000,80000000,1.5e7,1.5e8,60,50,90,60,90,80),each=8)
  
  for (i in 1:length(new)) {
    if (i < 49) {
      new[[i]] <- new[[i]] + scale_x_continuous(trans = "log10", limits = c(xmin[i], xmax[i])) + scale_y_continuous(trans = "log10", limits = c(ymin[i], ymax[i]))
    } else {
      new[[i]] <- new[[i]] + scale_x_continuous(trans = "log10", limits = c(xmin[i], xmax[i])) + scale_y_continuous(limits = c(ymin[i], ymax[i]))
    }
    
  }

  plots <- ggmatrix_gtable(ggmatrix(new, nrow=length(Ycols), ncol=length(Xcols), xAxisLabels = p$xAxisLabels[c(1:length(Xcols))], yAxisLabels = p$yAxisLabels[c(length(Xcols):(length(Xcols)+length(Ycols)-1))], labeller = label_wrap_gen(8), legend = c(2, 1), title = p$title) + theme(strip.background=element_rect(fill="white", color = "black"), legend.position = "bottom"))

  p <- ggarrange(plots)
  annotate_figure(p, fig.lab = paste0("Figure 6.2.", caption.num), fig.lab.pos = "bottom.right")
  
}

#Function to calculate geometric mean using log10 input - calculate mean of log10 values, then exponentiate back
geo_mean <- function(logX) {
  #n <- length(logX)
  geo_mean <- 10^mean(logX, na.rm=T)
  #geo_mean <- prod(x)^(1/n)
  return(geo_mean)
}

#Determine geometric mean ratio - function inputs are x and y on the log scale and regular scale, and the limit of detection for y
gmr <- function(logX, binY) {
  d <- data.frame(logX = logX, binY=binY)
  #Split the data based on whether y is above the LOD, or at or below the LOD
  #Get the geometric mean of the x (correlate) values of each of the 2 groups, and
  #determine the ratio
  gmr <- tryCatch(geo_mean(d[d$binY==1, "logX"])/geo_mean(d[d$binY==0, "logX"]), error = function(e) NA)
  gmr <- tryCatch(if (gmr > 10000) {
    gmr <- formatC(gmr, format = "e", digits = 2)
  } else {
    gmr <- round(gmr, 2)
  }, error=function(e) NA)
  #T-test on the log10-correlate value between the 2 groups (y above LOD, or y at/below LOD)
  #Do same t-test to get confidence interaval bounds, then back transform by raising the upper and lower
  #CI limits to the 10th power
  tst <- tryCatch(t.test(d[d$binY==1, "logX"], d[d$binY==0, "logX"]), error=function(e) NA)
  lower <- tryCatch(10^(tst)$conf.int[1], error=function(e) NA)
  upper <- tryCatch(10^(tst)$conf.int[2], error=function(e) NA)
  upper <- tryCatch(if (upper > 10000) {
    upper <- formatC(upper, format = "e", digits = 2)
  } else {
    upper <- round(upper, 2)
  }, error=function(e) NA)
  lower <- tryCatch(if (lower > 10000) {
    lower <- formatC(lower, format = "e", digits = 2)
  } else {
    lower <- round(lower, 2)
  }, error=function(e) NA)
  df <- data.frame(result = ifelse(!is.na(gmr) & !is.na(lower) & !is.na(upper), yes = paste0(gmr, " (", lower, ", ", upper, ")"), no = ifelse((is.na(lower) | is.na(upper)) & !is.na(gmr), yes = paste0(gmr, " (NA)"), no=NA)), row.names = NULL)
  return(df)
}
```



```{r, echo = FALSE}
D <- as.data.frame(readxl::read_xlsx("S:/BRB/Staff Folders/Allyson Mateja/NHP/data/allDat.xlsx"))
D$manuf <- factor(D$`Vaccine Candidate`, levels = c("Janssen Single Dose", "Janssen Two Dose", "Moderna", "Novavax", "Sanofi", "Placebo"), labels = c("Janssen", "Janssen", "Moderna", "Novavax", "Sanofi", "Placebo"))
D$`Challenge Day` <- factor(D$`Challenge Day`)
```


```{r, echo = FALSE}
cont.outcomes.log <- grep("log_VL|AUC", names(D))
cont.outcomes.reg <- grep("VL Day 2|AUC", names(D))
bin.outcomes <- grep("bin", names(D))

immune.markers.log <- grep("log_MN|log_PsVNA|log_MSD", names(D))
immune.markers.reg <- grep("(?<!log_)MN|(?<!log_)PsVNA|(?<!log_)MSD", names(D), perl=T)
```

\setcounter{table}{0}
\renewcommand{\thetable}{6.2.\arabic{table}}


# 6.2 Graphical Exploration of Virus-Specific Antibodies and Viral Load

## Virus-Specific Antibodies

### Plots

Scatter plots of pairs of virus-specific antibodies will be generated to determine how well they correlate. All arms (different vaccine products, different dilutions, and controls) will be included on the same figure with different colors for dose groups and different panels for vaccine product.  

```{r, echo = FALSE, warning=F, message=F,   fig.width=8, fig.height=10}
for (i in 1:length(unique(D$manuf))) {
  print(immune_markers_pairs(D[D$manuf == (unique(D$manuf)[i]), ], cols=immune.markers.reg, i))
  
}
```



\newpage

### Correlations

Simple statistics will be performed independent of the vaccine product. For example, the Spearman correlation between each pair of virus-specific antibodies will be calculated. Spearman (i.e., rank) correlations with confidence intervals will be generated; confidence intervals for the Spearman’s correlation coefficient are calculated by using the ranks in a Pearson’s correlation, which generally performs well (Caruso and Cliff, 1997). As there were not many ties, nonparametric bootstrap as a sensitivity analysis was not done. 

```{r, echo = FALSE}
combs <- combn(immune.markers.log, 2)
outcomes <- matrix(NA, nrow=length(immune.markers.log), ncol = length(immune.markers.log))

for (i in 1:ncol(combs)) {
  outcomes[which(immune.markers.log == combs[2, i]), which(immune.markers.log == combs[1, i])] <- cor_return(D[ ,combs[1, i]], D[ ,combs[2, i]])[1, 1]
}

outcomes <- as.data.frame(outcomes)
names(outcomes) <- gsub("_", " ", names(D)[immune.markers.log])
names(outcomes) <- gsub(".", " ", names(outcomes), fixed=T)

row.names(outcomes) <- gsub("_", " ", names(D)[immune.markers.log])
row.names(outcomes) <- gsub(".", " ", row.names(outcomes), fixed=T)

outcomes <- apply(outcomes, c(1, 2), function(x){ifelse(is.na(x), "-", x)})

kable(outcomes, row.names = TRUE, caption = "Spearman Correlations of Virus-Specific Antibodies", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "5em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```




\newpage

## Viral Load

### Plots

Pairwise plots of each virus-specific antibody with each viral load measurement will be generated. Data will be plotted with different colors for dose group and different panels for vaccine product.  

```{r, echo = FALSE,  warning = FALSE, message=FALSE, fig.width=8, fig.height=12}
for (i in 1:length(unique(D$manuf))) {
  print(markers_outcomes_pairs(D[D$manuf == (unique(D$manuf)[i]), ], Xcols=immune.markers.reg, Ycols=cont.outcomes.reg, i+length(unique(D$manuf))))
  
}
```

\newpage




### Correlations - Numeric Viral Load

Spearman's correlations of all numeric viral load measurements with each other will be calculated. 

```{r, echo = FALSE}
combs <- combn(cont.outcomes.log, 2)
outcomes <- matrix(NA, nrow=length(cont.outcomes.log), ncol = length(cont.outcomes.log))

for (i in 1:ncol(combs)) {
  outcomes[which(cont.outcomes.log == combs[2, i]), which(cont.outcomes.log == combs[1, i])] <- cor_return(D[ ,combs[1, i]], D[ ,combs[2, i]])[1, 1]
}

outcomes <- as.data.frame(outcomes)
names(outcomes) <- gsub("_", " ", names(D)[cont.outcomes.log])
names(outcomes) <- gsub(".", " ", names(outcomes), fixed=T)
#names(outcomes) <- str_wrap(names(outcomes), 20)
names(outcomes) <- gsub("PostChallenge", "Post Challenge", names(outcomes))

row.names(outcomes) <- gsub("_", " ", names(D)[cont.outcomes.log])
row.names(outcomes) <- gsub(".", " ", row.names(outcomes), fixed=T)

outcomes <- apply(outcomes, c(1, 2), function(x){ifelse(is.na(x), "-", x)})

kable(outcomes, row.names = TRUE, caption = "Spearman Correlations of Viral Load Outcomes", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(cont.outcomes.log)+1)), width = "2.5em") 
#%>% row_spec(0, angle = 90)
```

\newpage

For numeric viral load, Spearman correlations with each virus-specific Ab will be calculated. 

```{r, echo = FALSE}
combs2 <- expand.grid(cont.outcomes.log, immune.markers.log)

outcomes2 <- matrix(NA, ncol=length(immune.markers.log), nrow = length(cont.outcomes.log))


for (i in 1:nrow(combs2)) {
  outcomes2[which(cont.outcomes.log == combs2[i, 1]), which(immune.markers.log == combs2[i, 2])] <- cor_return(D[ ,combs2[i, 1]], D[ ,combs2[i, 2]])[1, 1]
}

outcomes2 <- as.data.frame(outcomes2)
names(outcomes2) <- gsub("_", " ", names(D)[immune.markers.log])
names(outcomes2) <- gsub(".", " ", names(outcomes2), fixed=T)

row.names(outcomes2) <- gsub("_", " ", names(D)[cont.outcomes.log])
row.names(outcomes2) <- gsub(".", " ", row.names(outcomes2), fixed=T)

kable(outcomes2, row.names = TRUE, caption = "Spearman Correlations of Virus-Specific Antibodies with Viral Load Measurements", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```

\newpage

### GMR - Binary Viral Load 

For binary viral load, the geometric mean ratio (GMR) as the geometric mean of the virus-specific antibodies from those animals with a positive signal for the viral load measurement (e.g., viral sub-genomic RNA for nasal swab at day 2 post-challenge) over the geometric mean of the virus-specific antibodies from those animals without a positive signal will be calculated. The confidence interval and tests on the GMR will be calculated by t-test on the difference in means of the log10 virus-specific antibody values, and back-transforming it to GMRs. NA is indicated when the binary responses are the same. 


```{r, echo = FALSE}
combs3 <- expand.grid(immune.markers.log, bin.outcomes)

outcomes3 <- matrix(NA, nrow=length(immune.markers.log), ncol = length(bin.outcomes))


for (i in 1:nrow(combs3)) {
  outcomes3[which(immune.markers.log == combs3[i, 1]), which(bin.outcomes == combs3[i, 2])] <- gmr(D[ ,combs3[i, 1]], D[ ,combs3[i, 2]])[1, 1]
}

outcomes3 <- as.data.frame(outcomes3)
names(outcomes3) <- gsub("_", " ", names(D)[bin.outcomes])
names(outcomes3) <- gsub(".", " ", names(outcomes3), fixed=T)

row.names(outcomes3) <- gsub("_", " ", names(D)[immune.markers.log])
row.names(outcomes3) <- gsub(".", " ", row.names(outcomes3), fixed=T)

kable(outcomes3, row.names = TRUE, caption = "Geometric Mean Ratios of Virus-Specific Antibodies with Binary Viral Load Measurements", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "5em") %>% column_spec(c(2:(length(bin.outcomes)+1)), width = "5em") %>% footnote(general = "NA is indicated when there is only 1 negative signal and so a 95% CI can not be calculated.")
```


