---
title: "CoP SAP 6.5 and 6.6 Appendix"
author: "Allyson Mateja"
date: "`r Sys.Date()`"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)
library(boot)
library(dplyr)
library(readxl)
library(saws)
```


```{r, echo = FALSE}
#Functions 


#Function to calculate adjusted R2
#Input y values, predicted y values, and number of parameters (including intercept) in the model
adj_rsquared <- function(y, yhat, p){
  #Caculate r squared using normal formula
  r_sq <- 1 - (sum((y-yhat)^2)/sum((y-mean(y))^2))
  n <- length(y) #Get sample size
  #Adjust R squared by sample size and number of model parameters
  adj_r_sq <- 1-(((n-1)*(1-r_sq))/(n-p))
  return(adj_r_sq)
}

#Function to calculate adjusted coefficient of discrimination
#Input y values, predicted y values, and number of parameters (including intercept) in the model
adj_cod <- function(y, yhat, p) {
  #Calculate CoD using normal formula
  if (length(yhat[y==0]) > 0 & length(yhat[y==1]) > 0) {
    cod<-mean(yhat[y==1], na.rm=T) -mean(yhat[y==0], na.rm=T)
  } else if (length(yhat[y==0]) > 0 & length(yhat[y==1]) == 0) {
    cod<-mean(yhat[y==0], na.rm=T)
  } else if (length(yhat[y==0]) == 0 & length(yhat[y==1]) > 0) {
    cod<-mean(yhat[y==1], na.rm=T)
  }
  
  n <- length(y)
  #Adjust CoD by sample size and number of model parameters
  adj_cod <- 1-(((n-1)*(1-cod))/(n-p))
  return(adj_cod)
}

#NOTE: Treating dose score as continuous
#We may need to adjust the dose score value, and calibrate it to the Beta values from 6.3.1
#This may need to change

###6.5

#Function for bootstrap of PVE for continuous outcome
boot.pve_base_dose_adj_r_sq_lm <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  #Get PVE for bootstrap sample
  
  base_mod <- lm(logY ~ logX, data=df2)
  base_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_mod), p=base_mod$rank)
  
  base_dose_mod <- lm(logY ~ logX*dose_score, data=df2)
  base_dose_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_dose_mod), p=base_dose_mod$rank)
  
  #Calculate PVE
  pve <- 100*(base_r_sq/base_dose_r_sq)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
  
}


boot.pve_base_dose_adj_r_sq_lm_rsq <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  #Get PVE for bootstrap sample
  
  base_mod <- lm(logY ~ logX, data=df2)
  base_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_mod), p=base_mod$rank)
  
  base_dose_mod <- lm(logY ~ logX*dose_score, data=df2)
  base_dose_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_dose_mod), p=base_dose_mod$rank)
  
  #Calculate PVE
  pve <- 100*(base_r_sq/base_dose_r_sq)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(c(base_r_sq, base_dose_r_sq))
  
}



boot.pve_base_dose_adj_r_sq_exp <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  #Get PVE for bootstrap sample
  
  base_mod <- glm(logY ~ logX, data=df2, family=gaussian(link="log"))
  base_r_sq <- adj_rsquared(y=df2$logY, yhat=exp(predict(base_mod)), p=base_mod$rank)
    
  base_dose_mod <- glm(logY ~ logX*dose_score, data=df2, family=gaussian(link="log"))
  base_dose_r_sq <- adj_rsquared(y=df2$logY, yhat=exp(predict(base_dose_mod)), p=base_dose_mod$rank)
    
  #Calculate PVE
  pve <- 100*(base_r_sq/base_dose_r_sq)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
  
}


#Function to determine PVE and 95% bootstrap confidence interval for continuous outcome
return_all_pve_base_dose_adj_r_sq <- function(logX, logY, model, dose_score = D_doses$dose_score, R=5000, seed=28414987) {
  #set.seed(seed)
  #Create data frame from input variables
  df <- data.frame(logX=logX, logY=logY, dose_score=dose_score)
  
  df <- df[complete.cases(df), ]
  #df$dose_score <- factor(df$dose_score)
  
  #Using full data, fit base and base+dose models, get corresponding R squared
  
  if (model == "lm") {
    base_mod <- lm(logY ~ logX, data=df)
    base_r_sq <- adj_rsquared(y=df$logY, yhat=predict(base_mod), p=base_mod$rank)
  
    base_dose_mod <- lm(logY ~ logX*dose_score, data=df)
    base_dose_r_sq <- adj_rsquared(y=df$logY, yhat=predict(base_dose_mod), p=base_dose_mod$rank)
    
    #Calculate PVE
    pve <- 100*(base_r_sq/base_dose_r_sq)
    pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  
    #Get bootstrap samples
    boot_samp <- boot(df, boot.pve_base_dose_adj_r_sq_lm, R=R)
    ci <- boot.ci(boot_samp, type = "bca")$bca
  
    #Get quantiles of bootstrap sample corresponding to 95% confidence interval
    lower <- ci[4]
    upper <- ci[5]
  
    lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
    upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
  
    
    
  } else if (model == "exp") {
    
    base_mod <- glm(logY ~ logX, data=df, family=gaussian(link="log"))
    base_r_sq <- adj_rsquared(y=df$logY, yhat=exp(predict(base_mod)), p=base_mod$rank)
    
    base_dose_mod <- glm(logY ~ logX*dose_score, data=df, family=gaussian(link="log"))
    base_dose_r_sq <- adj_rsquared(y=df$logY, yhat=exp(predict(base_dose_mod)), p=base_dose_mod$rank)
    
    
    #Calculate PVE
    pve <- 100*(base_r_sq/base_dose_r_sq)
    pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  
    #Get bootstrap samples
    boot_samp <- boot(df, boot.pve_base_dose_adj_r_sq_exp, R=R)
    ci <- boot.ci(boot_samp, type = "bca")$bca
  
    #Get quantiles of bootstrap sample corresponding to 95% confidence interval
    lower <- ci[4]
    upper <- ci[5]
  
    lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
    upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
  
    
  }
  
  #Make dataframe of results to return
  boot_return <- data.frame(result = paste0(round(pve, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  return(boot_return)
}

#Function for bootstrap of PVE for binary outcome
boot.pve_base_dose_cod <- function(df, i) {
  #Select bootstrap sample with replcaement
  df2 <- df[i, ]
  
  #Calculate PVE from bootstrap sample
  base_mod <- glm(binY ~ logX, data=df2, family="binomial")
  base_cod <- adj_cod(y=df2$binY, yhat=predict(base_mod, type="response"), p=base_mod$rank)
  
  base_dose_mod <- glm(binY ~ logX*dose_score, data=df2, family="binomial")
  base_dose_cod <- adj_cod(y=df2$binY, yhat=predict(base_dose_mod, type="response"), p=base_dose_mod$rank)
  
  pve <- 100*(base_cod/base_dose_cod)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
}



#Function to determine PVE and 95% bootstrap confidence interval for binary outcome
return_all_pve_base_dose_cod <- function(logX, binY, dose_score = D_doses$dose_score, R=5000, seed=28414987) {
  #Get PVE
  set.seed(seed)
  #Create data frame from input variables, add in binary variable for above/below limit of detection
  df <- data.frame(logX=logX, binY=binY, dose_score=dose_score)
  df <- df[complete.cases(df), ]
  #df$dose_score <- factor(df$dose_score)
  #Using full data, fit base and base+dose models, get corresponding CoD
  base_mod <- glm(binY ~ logX, data=df, family="binomial")
  base_cod <- adj_cod(y=df$binY, yhat=predict(base_mod, type="response"), p=base_mod$rank)
  
  base_dose_mod <- glm(binY ~ logX*dose_score, data=df, family="binomial")
  base_dose_cod <- adj_cod(y=df$binY, yhat=predict(base_dose_mod, type="response"), p=base_dose_mod$rank)
  
  #Calculate PVE
  pve <- 100*(base_cod/base_dose_cod)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  
  
  #Get bootstrap samples
  boot_samp <- boot(df, boot.pve_base_dose_cod, R=R)
  ci <- tryCatch(boot.ci(boot_samp, type = "bca")$bca, error=function(e) boot.ci(boot_samp, type = "perc")$perc)
  
  #Get quantiles of bootstrap sample corresponding to 95% confidence interval
  lower <- ci[4]
  upper <- ci[5]
  
  lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
  upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
  
  
  #Make dataframe of results to return
  boot_return <- data.frame(result = paste0(round(pve, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  return(boot_return)
}



### 6.6

#NOTE: Treating dose score as continuous
#We may need to adjust the dose score value, and calibrate it to the Beta values from 6.3.1
#This may need to change

#Function for bootstrap of PVE for continuous outcome
boot.pve_dose_vacc_adj_r_sq_lm <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  #Get PVE for bootstrap sample
  base_dose_mod <- lm(logY ~ logX*dose_score, data=df2)
  base_dose_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_dose_mod), p=base_dose_mod$rank)
  
  base_dose_vacc_mod <- lm(logY ~ logX*dose_score*manuf, data=df2)
  base_dose_vacc_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_dose_vacc_mod), p=base_dose_vacc_mod$rank)
  
  #Calculate PVE
  pve <- 100*(base_dose_r_sq/base_dose_vacc_r_sq)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
  
}

boot.pve_dose_vacc_adj_r_sq_exp <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  #Get PVE for bootstrap sample
  base_dose_mod <- glm(logY ~ logX*dose_score, data=df2, family=gaussian(link="log"))
  base_dose_r_sq <- adj_rsquared(y=df2$logY, yhat=exp(predict(base_dose_mod)), p=base_dose_mod$rank)
    
  base_dose_vacc_mod <- glm(logY ~ logX*dose_score*manuf, data=df2, family=gaussian(link="log"))
  base_dose_vacc_r_sq <- adj_rsquared(y=df2$logY, yhat=exp(predict(base_dose_vacc_mod)), p=base_dose_vacc_mod$rank)
    
  #Calculate PVE
  pve <- 100*(base_dose_r_sq/base_dose_vacc_r_sq)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
  
}

#Function to determine PVE and 95% bootstrap confidence interval for continuous outcome
return_all_pve_dose_vacc_adj_r_sq <- function(logX, logY, model, dose_score = D_doses$dose_score, manuf=D_doses$manuf, R=5000) {
 
  #Create data frame from input variables
  df <- data.frame(logX=logX, logY=logY, dose_score=dose_score, manuf=manuf)
  df <- df[complete.cases(df), ]
  
  #df$dose_score <- factor(df$dose_score)
  
  
  if (model == "lm") {
    
    base_dose_mod <- lm(logY ~ logX*dose_score, data=df)
    base_dose_r_sq <- adj_rsquared(y=df$logY, yhat=predict(base_dose_mod), p=base_dose_mod$rank)
    
    base_dose_vacc_mod <- lm(logY ~ logX*dose_score*manuf, data=df)
    base_dose_vacc_r_sq <- adj_rsquared(y=df$logY, yhat=predict(base_dose_vacc_mod), p=base_dose_vacc_mod$rank)
    
    #Calculate PVE
    pve <- 100*(base_dose_r_sq/base_dose_vacc_r_sq)
    pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  
    #Get bootstrap samples
    boot_samp <- boot(df, boot.pve_dose_vacc_adj_r_sq_lm, R=R)
    ci <- boot.ci(boot_samp, type = "bca")$bca
  
    #Get quantiles of bootstrap sample corresponding to 95% confidence interval
    lower <- ci[4]
    upper <- ci[5]
    
    lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
    upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
    
  } else if (model == "exp") {
    
    base_dose_mod <- glm(logY ~ logX*dose_score, data=df, family=gaussian(link="log"))
    base_dose_r_sq <- adj_rsquared(y=df$logY, yhat=exp(predict(base_dose_mod)), p=base_dose_mod$rank)
    
    base_dose_vacc_mod <- glm(logY ~ logX*dose_score*manuf, data=df, family=gaussian(link="log"))
    base_dose_vacc_r_sq <- adj_rsquared(y=df$logY, yhat=exp(predict(base_dose_vacc_mod)), p=base_dose_vacc_mod$rank)
    
    
    #Calculate PVE
    pve <- 100*(base_dose_r_sq/base_dose_vacc_r_sq)
    pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  
    #Get bootstrap samples
    boot_samp <- boot(df, boot.pve_dose_vacc_adj_r_sq_exp, R=R)
    ci <- boot.ci(boot_samp, type = "bca")$bca
  
    #Get quantiles of bootstrap sample corresponding to 95% confidence interval
    lower <- ci[4]
    upper <- ci[5]
  
    lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
    upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
  
    
  }
  
  #Make dataframe of results to return
  boot_return <- data.frame(result = paste0(round(pve, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  return(boot_return)
}



#Function for bootstrap of PVE for binary outcome
boot.pve_dose_vacc_cod <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  
  #Calculate PVE from bootstrap sample
  base_dose_mod <- glm(binY ~ logX*dose_score, data=df2, family="binomial")
  base_dose_cod <- adj_cod(y=df2$binY, yhat=predict(base_dose_mod, type="response"), p=base_dose_mod$rank)
  
  base_dose_vacc_mod <- glm(binY ~ logX*dose_score*manuf, data=df2, family="binomial")
  base_dose_vacc_cod <- adj_cod(y=df2$binY, yhat=predict(base_dose_vacc_mod, type="response"), p=base_dose_vacc_mod$rank)
  
  pve <- 100*(base_dose_cod/base_dose_vacc_cod)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
}



#Function to determine PVE and 95% bootstrap confidence interval for binary outcome
return_all_pve_dose_vacc_cod <- function(logX, binY, dose_score = D_doses$dose_score, manuf=D_doses$manuf, R=5000) {
  
  #Create data frame from input variables, add in binary variable for above/below limit of detection
  df <- data.frame(logX=logX, binY=binY, dose_score=dose_score, manuf=manuf)
  df <- df[complete.cases(df), ]
  
  #df$dose_score <- factor(df$dose_score)
  
  #Using full data, fit base+dose and base+dose+vacc models, get corresponding CoD
  base_dose_mod <- glm(binY ~ logX*dose_score, data=df, family="binomial")
  base_dose_cod <- adj_cod(y=df$binY, yhat=predict(base_dose_mod, type="response"), p=base_dose_mod$rank)
  
  base_dose_vacc_mod <- glm(binY ~ logX*dose_score*manuf, data=df, family="binomial")
  base_dose_vacc_cod <- adj_cod(y=df$binY, yhat=predict(base_dose_vacc_mod, type="response"), p=base_dose_vacc_mod$rank)
  
  #Get PVE
  pve <- 100*(base_dose_cod/base_dose_vacc_cod)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  #Get bootstrap samples
  boot_samp <- boot(df, boot.pve_dose_vacc_cod, R=R)
  ci <- tryCatch(boot.ci(boot_samp, type = "bca")$bca, error=function(e) boot.ci(boot_samp, type = "perc")$perc)
  
  #Get quantiles of bootstrap sample corresponding to 95% confidence interval
  lower <- ci[4]
  upper <- ci[5]
  
  lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
  upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
  
  #Make dataframe of results to return
  boot_return <- data.frame(result = paste0(round(pve, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  return(boot_return)
}





## Additional analysis
## Compare base to base+dose+vacc

boot.pve_base_vacc_adj_r_sq_lm <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  #Get PVE for bootstrap sample
  base_mod <- lm(logY ~ logX, data=df2)
  base_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_mod), p=base_mod$rank)
  
  base_dose_vacc_mod <- lm(logY ~ logX*dose_score*manuf, data=df2)
  base_dose_vacc_r_sq <- adj_rsquared(y=df2$logY, yhat=predict(base_dose_vacc_mod), p=base_dose_vacc_mod$rank)
  
  #Calculate PVE
  pve <- 100*(base_r_sq/base_dose_vacc_r_sq)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
  
}

boot.pve_base_vacc_adj_r_sq_exp <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  #Get PVE for bootstrap sample
  base_mod <- glm(logY ~ logX, data=df2, family=gaussian(link="log"))
  base_r_sq <- adj_rsquared(y=df2$logY, yhat=exp(predict(base_mod)), p=base_mod$rank)
    
  base_dose_vacc_mod <- glm(logY ~ logX*dose_score*manuf, data=df2, family=gaussian(link="log"))
  base_dose_vacc_r_sq <- adj_rsquared(y=df2$logY, yhat=exp(predict(base_dose_vacc_mod)), p=base_dose_vacc_mod$rank)
    
  #Calculate PVE
  pve <- 100*(base_r_sq/base_dose_vacc_r_sq)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
  
}

#Function to determine PVE and 95% bootstrap confidence interval for continuous outcome
return_all_pve_base_vacc_adj_r_sq <- function(logX, logY, model, dose_score = D_doses$dose_score, manuf=D_doses$manuf, R=5000) {
 
  #Create data frame from input variables
  df <- data.frame(logX=logX, logY=logY, dose_score=dose_score, manuf=manuf)
  df <- df[complete.cases(df), ]
  
  #df$dose_score <- factor(df$dose_score)
  
  
  if (model == "lm") {
    
    base_mod <- lm(logY ~ logX, data=df)
    base_r_sq <- adj_rsquared(y=df$logY, yhat=predict(base_mod), p=base_mod$rank)
    
    base_dose_vacc_mod <- lm(logY ~ logX*dose_score*manuf, data=df)
    base_dose_vacc_r_sq <- adj_rsquared(y=df$logY, yhat=predict(base_dose_vacc_mod), p=base_dose_vacc_mod$rank)
    
    #Calculate PVE
    pve <- 100*(base_r_sq/base_dose_vacc_r_sq)
    pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  
    #Get bootstrap samples
    boot_samp <- boot(df, boot.pve_base_vacc_adj_r_sq_lm, R=R)
    ci <- tryCatch(boot.ci(boot_samp, type = "bca")$bca, error=function(e) boot.ci(boot_samp, type = "perc")$perc)
  
    #Get quantiles of bootstrap sample corresponding to 95% confidence interval
    lower <- ci[4]
    upper <- ci[5]
    
    lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
    upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
    
  } else if (model == "exp") {
    
    base_mod <- glm(logY ~ logX, data=df, family=gaussian(link="log"))
    base_r_sq <- adj_rsquared(y=df$logY, yhat=exp(predict(base_mod)), p=base_mod$rank)
    
    base_dose_vacc_mod <- glm(logY ~ logX*dose_score*manuf, data=df, family=gaussian(link="log"))
    base_dose_vacc_r_sq <- adj_rsquared(y=df$logY, yhat=exp(predict(base_dose_vacc_mod)), p=base_dose_vacc_mod$rank)
    
    
    #Calculate PVE
    pve <- 100*(base_r_sq/base_dose_vacc_r_sq)
    pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  
    #Get bootstrap samples
    boot_samp <- boot(df, boot.pve_base_vacc_adj_r_sq_exp, R=R)
    ci <- tryCatch(boot.ci(boot_samp, type = "bca")$bca, error=function(e) boot.ci(boot_samp, type = "perc")$perc)
  
    #Get quantiles of bootstrap sample corresponding to 95% confidence interval
    lower <- ci[4]
    upper <- ci[5]
  
    lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
    upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
  
    
  }
  
  #Make dataframe of results to return
  boot_return <- data.frame(result = paste0(round(pve, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  return(boot_return)
}



#Function for bootstrap of PVE for binary outcome
boot.pve_base_vacc_cod <- function(df, i) {
  #Select bootstrap sample with replacement
  df2 <- df[i, ]
  
  #Calculate PVE from bootstrap sample
  base_mod <- glm(binY ~ logX, data=df2, family="binomial")
  base_cod <- adj_cod(y=df2$binY, yhat=predict(base_mod, type="response"), p=base_mod$rank)
  
  base_dose_vacc_mod <- glm(binY ~ logX*dose_score*manuf, data=df2, family="binomial")
  base_dose_vacc_cod <- adj_cod(y=df2$binY, yhat=predict(base_dose_vacc_mod, type="response"), p=base_dose_vacc_mod$rank)
  
  pve <- 100*(base_cod/base_dose_vacc_cod)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  return(pve)
}



#Function to determine PVE and 95% bootstrap confidence interval for binary outcome
return_all_pve_base_vacc_cod <- function(logX, binY, dose_score = D_doses$dose_score, manuf=D_doses$manuf, R=5000) {
  
  #Create data frame from input variables, add in binary variable for above/below limit of detection
  df <- data.frame(logX=logX, binY=binY, dose_score=dose_score, manuf=manuf)
  df <- df[complete.cases(df), ]
  
  #df$dose_score <- factor(df$dose_score)
  
  #Using full data, fit base+dose and base+dose+vacc models, get corresponding CoD
  base_mod <- glm(binY ~ logX, data=df, family="binomial")
  base_cod <- adj_cod(y=df$binY, yhat=predict(base_mod, type="response"), p=base_mod$rank)
  
  base_dose_vacc_mod <- glm(binY ~ logX*dose_score*manuf, data=df, family="binomial")
  base_dose_vacc_cod <- adj_cod(y=df$binY, yhat=predict(base_dose_vacc_mod, type="response"), p=base_dose_vacc_mod$rank)
  
  #Get PVE
  pve <- 100*(base_cod/base_dose_vacc_cod)
  pve <- ifelse(pve < 0, 0, ifelse(pve > 100, 100, pve))
  #Get bootstrap samples
  boot_samp <- boot(df, boot.pve_base_vacc_cod, R=R)
  ci <- tryCatch(boot.ci(boot_samp, type = "bca")$bca, error=function(e) boot.ci(boot_samp, type = "perc")$perc)
  
  #Get quantiles of bootstrap sample corresponding to 95% confidence interval
  lower <- ci[4]
  upper <- ci[5]
  
  lower <- ifelse(lower < 0, 0, ifelse(lower > 100, 100, lower))
  upper <- ifelse(upper < 0, 0, ifelse(upper > 100, 100, upper))
  
  #Make dataframe of results to return
  boot_return <- data.frame(result = paste0(round(pve, 2), " (", round(lower, 2), ", ", round(upper, 2), ")"), row.names = NULL)
  return(boot_return)
}





```


```{r, echo = FALSE}
D <- as.data.frame(readxl::read_xlsx("H:/Projects/Mike Fay/NHP/Final Data/allDat.xlsx"))
D$manuf <- factor(D$`Vaccine Candidate`, levels = c("Janssen Single Dose", "Janssen Two Dose", "Moderna", "Novavax", "Sanofi", "Placebo"), labels = c("Janssen", "Janssen", "Moderna", "Novavax", "Sanofi", "Placebo"))
D$`Challenge Day` <- factor(D$`Challenge Day`)
```



```{r, echo = FALSE}
#Input doses from vaccine manufacturers
doses <- data.frame(Subgroup = c("1A", "1B", "1C", "1D", "1E", "2A", "2B", "2C", "2D", "2E", "2F", "3A", "3B", "3C", "3D", "3E", "4A", "4B", "4C", "4D", "4E"), dose = c(30, 10, 3, 1, 0.3, 5*10^10, 1.25*10^10, 5*10^9, 2*10^9, 5*10^10, 5*10^9, 5, 1, 0.2, 0.04, 0.008, 15, 7.5, 3.75, 1.88, 0.94), min_dose = c(rep(0.3, 5), rep(2*10^9, 6), rep(0.008, 5), rep(0.94, 5)))

doses$Subgroup <- factor(doses$Subgroup)
```

```{r, echo = FALSE}
#Merge actual dose information with data
D_doses <- merge(D, doses, by = "Subgroup", all.x=T)

#Calculate dose score
D_doses$dose_score <- log10((10*D_doses$dose)/D_doses$min_dose)
D_doses$dose_score <- ifelse(D_doses$Subgroup == "5A", yes=0, no=D_doses$dose_score)
#Calculate second dose score for single dose groups, by dividing original dose in half
D_doses$dose_score_d1_half <- ifelse(D_doses$Subgroup %in% c("2A", "2B", "2C", "2D"), yes = log10((10*(D_doses$dose/2))/(D_doses$min_dose/2)), no = ifelse(D_doses$Subgroup %in% c("2E", "2F"), yes = log10((10*(D_doses$dose))/(D_doses$min_dose/2)), no = D_doses$dose_score))
D_doses$dose_score_double <- ifelse(D_doses$Subgroup %in% c("2E", "2F"), yes = log10((10*(D_doses$dose))/min(D_doses[D_doses$Subgroup %in% c("2E", "2F"), ]$dose)), no = ifelse(D_doses$Subgroup %in% c("2A", "2B", "2C", "2D"), yes = NA, D_doses$dose_score))


D_doses$dose_unit <- ifelse(D_doses$manuf %in% c("Moderna", "Novavax", "Sanofi"), yes = "ug", no = ifelse(D_doses$manuf == "Janssen", yes = "Viral\nParticles", no = ""))

D_doses$dose_char <- paste0(D_doses$dose, " ", D_doses$dose_unit)
D_doses$dose_char <- ifelse(D_doses$manuf == "Placebo", "0 ug", D_doses$dose_char)

D_doses$dose_char <- factor(D_doses$dose_char, levels = c("0 ug", "0.008 ug", "0.04 ug", "0.2 ug", "0.3 ug", "0.94 ug", "1 ug", "1.88 ug", "3 ug", "3.75 ug", "5 ug", "7.5 ug", "10 ug", "15 ug", "30 ug", "2e+09 Viral\nParticles", "5e+09 Viral\nParticles", "1.25e+10 Viral\nParticles", "5e+10 Viral\nParticles"))


```

```{r, echo = FALSE}
cont.outcomes.log <- grep("log_VL|AUC", names(D_doses))
cont.outcomes.reg <- grep("VL Day 2|AUC", names(D_doses))
bin.outcomes <- grep("bin_", names(D_doses))

immune.markers.log <- grep("log_MN|log_PsVNA|log_MSD", names(D_doses))
immune.markers.reg <- grep("(?<!log_)MN|(?<!log_)PsVNA|(?<!log_)MSD", names(D_doses), perl=T)
```



```{r, echo = FALSE}
cv_outcomes <- readxl::read_xlsx("C:/Users/matejaam/Desktop/cv_outcomes.xlsx")
cv_outcomes_bin <- readxl::read_xlsx("C:/Users/matejaam/Desktop/cv_outcomes_bin.xlsx")
```

```{r, echo = FALSE}
D_doses$dose_score_binned <- factor(if_else(D_doses$dose_score_d1_half > 1 & D_doses$dose_score_d1_half <= 3, cut(D_doses$dose_score_d1_half, breaks = seq(1, 3, 0.5)), if_else(D_doses$dose_score_d1_half > 3, ">3", factor(D_doses$dose_score_d1_half))), levels = c("0", "1", "(1,1.5]", "(1.5,2]", "(2,2.5]", "(2.5,3]", ">3"))

D_doses$dose_score_binned_large <- factor(if_else(D_doses$dose_score_d1_half > 1 & D_doses$dose_score_d1_half <= 2, "(1, 2]", if_else(D_doses$dose_score_d1_half > 2, ">2", factor(D_doses$dose_score_d1_half))), levels = c("0", "1", "(1, 2]", ">2"))
```

\setcounter{table}{0}
\renewcommand{\thetable}{A6.5.\arabic{table}}

# 6.5 Appendix

As an appendix to section 6.5, we repeat the analysis using binned dose scores. The single-dose halved dose scores are binned as follows: 0, 1, (1, 1.5], (1.5, 2], (2, 2.5], (2.5, 3], > 3. This is being done in an attempt to improve the performance of the bootstrap confidence intervals. 

## Numeric viral load

```{r, echo = FALSE, cache=T, warning=F, message=F}
combs2 <- expand.grid(immune.markers.log, cont.outcomes.log)
combs2$immune <- names(D_doses)[combs2[ ,1]]

combs2$immune <- gsub("_", " ", combs2$immune)
combs2$immune <- gsub(".", " ", combs2$immune, fixed=T)

combs2$outcome <- names(D_doses)[combs2[ ,2]]
combs2$outcome <- gsub("_", " ", combs2$outcome)
combs2$outcome <- gsub(".", " ", combs2$outcome, fixed=T)

combs2 <- merge(combs2, cv_outcomes, by = c("immune", "outcome"))

outcomes2 <- matrix(NA, ncol=length(immune.markers.log), nrow = length(cont.outcomes.log))


for (i in 1:nrow(combs2)) {
  outcomes2[which(cont.outcomes.log == combs2[i, "Var2"]), which(immune.markers.log == combs2[i, "Var1"])] <- return_all_pve_base_dose_adj_r_sq(logX=D_doses[ ,combs2[i, "Var1"]], logY=D_doses[ ,combs2[i, "Var2"]], dose_score = D_doses$dose_score_binned, model = combs2[i, "best.fit"], R=5000)[1, 1]
}

outcomes2 <- as.data.frame(outcomes2)
row.names(outcomes2) <- gsub("_", " ", names(D_doses)[cont.outcomes.log])
row.names(outcomes2) <- gsub(".", " ", row.names(outcomes2), fixed=T)

names(outcomes2) <- gsub("_", " ", names(D_doses)[immune.markers.log])
names(outcomes2) <- gsub(".", " ", names(outcomes2), fixed=T)
```

```{r, echo = FALSE, warning = FALSE, message=FALSE}

kable(outcomes2, row.names = TRUE, caption = "Percent of Variance Explained by Immune Virus-Specific Antibodies - Numeric Viral Load with Binned Dose Score", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```

## Binary Viral Load


```{r, echo = FALSE, warning=FALSE, message=FALSE, cache=T}
combs3 <- expand.grid(immune.markers.log, bin.outcomes)

outcomes3 <- matrix(NA, ncol=length(immune.markers.log), nrow = length(bin.outcomes))


for (i in 1:nrow(combs3)) {
  outcomes3[which(bin.outcomes == combs3[i, 2]), which(immune.markers.log == combs3[i, 1])] <- return_all_pve_base_dose_cod(logX=D_doses[ ,combs3[i, 1]], binY=D_doses[ ,combs3[i, 2]], dose_score = D_doses$dose_score_binned, R=5000)[1, 1]
}

outcomes3 <- as.data.frame(outcomes3)
row.names(outcomes3) <- gsub("_", " ", names(D_doses)[bin.outcomes])
row.names(outcomes3) <- gsub(".", " ", row.names(outcomes3), fixed=T)

names(outcomes3) <- gsub("_", " ", names(D_doses)[immune.markers.log]) 
names(outcomes3) <- gsub(".", " ", names(outcomes3), fixed=T)

```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
kable(outcomes3, row.names = TRUE, caption = "Percent of Variance Explained by Immune Virus Specific Antibodies - Binary Viral Load with Binned Dose Score", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```


\setcounter{table}{0}
\renewcommand{\thetable}{A6.6.\arabic{table}}


# 6.6 Appendix

SAP section 6.6 is also repeated using larger bins of the single dose halved dose scores: 0, 1, (1, 2], >2. This was done to increase the amount of data in each bin, after splitting by vaccine.  


## Numeric Viral Load

```{r, echo = FALSE, cache=T, warning=F, message=F}
outcomes4 <- matrix(NA, ncol=length(immune.markers.log), nrow = length(cont.outcomes.log))


for (i in 1:nrow(combs2)) {
  outcomes4[which(cont.outcomes.log == combs2[i, "Var2"]), which(immune.markers.log == combs2[i, "Var1"])] <- return_all_pve_dose_vacc_adj_r_sq(logX=D_doses[ ,combs2[i, "Var1"]], logY=D_doses[ ,combs2[i, "Var2"]], dose_score = D_doses$dose_score_binned_large, model=combs2[i, "best.fit"], R=5000)[1, 1]
}

outcomes4 <- as.data.frame(outcomes4)
row.names(outcomes4) <- gsub("_", " ", names(D_doses)[cont.outcomes.log])
row.names(outcomes4) <- gsub(".", " ", row.names(outcomes4), fixed=T)

names(outcomes4) <- gsub("_", " ", names(D_doses)[immune.markers.log])
names(outcomes4) <- gsub(".", " ", names(outcomes4), fixed=T)

kable(outcomes4, row.names = TRUE, caption = "Differential Vaccine Effect - Numeric Viral Load with Binned Dose Score", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```



## Binary Viral Load

```{r, echo = FALSE, warning=FALSE, message=FALSE, cache=T}
outcomes5 <- matrix(NA, ncol=length(immune.markers.log), nrow = length(bin.outcomes))


for (i in 1:nrow(combs3)) {
  outcomes5[which(bin.outcomes == combs3[i, 2]), which(immune.markers.log == combs3[i, 1])] <- return_all_pve_dose_vacc_cod(logX=D_doses[ ,combs3[i, 1]], binY=D_doses[ ,combs3[i, 2]],dose_score = D_doses$dose_score_binned_large, R=5000)[1, 1]
}

outcomes5 <- as.data.frame(outcomes5)
row.names(outcomes5) <- gsub("_", " ", names(D_doses)[bin.outcomes])
row.names(outcomes5) <- gsub(".", " ", row.names(outcomes5), fixed=T)

names(outcomes5) <- gsub("_", " ", names(D_doses)[immune.markers.log])
names(outcomes5) <- gsub(".", " ", names(outcomes5), fixed=T)

kable(outcomes5, row.names = TRUE, caption = "Differential Vaccine Effect - Binary Viral Load with Binned Dose Score", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```





# Additional Analysis - Base+Dose+Vacc vs. Base

As an additional analysis for the appendix, we compare the base model to the base+dose+vacc model. Details of the base model are explained in the SAP section 6.5. The base+dose+vacc model used here is slightly different than that described in SAP section 6.6. In this model, the single dose halved dose score is used as a numeric value, with a separate effect for each vaccine manufacturer. Linear models for numeric viral load (e.g., Section 6.4.1) or logistic models for binary viral load (more complicated models, regardless of what was selected by the cross-validation procedure in section 6.4.2, are not used) will be employed. The percentage of variance (PVE) is 100 times the adjusted R square value from the base model over the adjusted R squared values from the base+dose+vacc model. Confidence intervals on the PVE are calculated by using bias-adjusted nonparametric bootstrap methods (using 5000 replicates). 


## Numeric Viral Load

```{r, echo = FALSE, cache=T, warning=F, message=F}
outcomes6 <- matrix(NA, ncol=length(immune.markers.log), nrow = length(cont.outcomes.log))


for (i in 1:nrow(combs2)) {
  outcomes6[which(cont.outcomes.log == combs2[i, "Var2"]), which(immune.markers.log == combs2[i, "Var1"])] <- return_all_pve_base_vacc_adj_r_sq(logX=D_doses[ ,combs2[i, "Var1"]], logY=D_doses[ ,combs2[i, "Var2"]], dose_score = D_doses$dose_score_d1_half, model=combs2[i, "best.fit"], R=5000)[1, 1]
}

outcomes6 <- as.data.frame(outcomes6)
row.names(outcomes6) <- gsub("_", " ", names(D_doses)[cont.outcomes.log])
row.names(outcomes6) <- gsub(".", " ", row.names(outcomes6), fixed=T)

names(outcomes6) <- gsub("_", " ", names(D_doses)[immune.markers.log])
names(outcomes6) <- gsub(".", " ", names(outcomes6), fixed=T)

kable(outcomes6, row.names = TRUE, caption = "Compare Base to Base, Dose, Vacc - Numeric Viral Load with Binned Dose Score", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```



## Binary Viral Load

```{r, echo = FALSE, warning=FALSE, message=FALSE, cache=T}
outcomes7 <- matrix(NA, ncol=length(immune.markers.log), nrow = length(bin.outcomes))


for (i in 1:nrow(combs3)) {
  outcomes7[which(bin.outcomes == combs3[i, 2]), which(immune.markers.log == combs3[i, 1])] <- return_all_pve_base_vacc_cod(logX=D_doses[ ,combs3[i, 1]], binY=D_doses[ ,combs3[i, 2]],dose_score = D_doses$dose_score_d1_half, R=5000)[1, 1]
}

outcomes7 <- as.data.frame(outcomes7)
row.names(outcomes7) <- gsub("_", " ", names(D_doses)[bin.outcomes])
row.names(outcomes7) <- gsub(".", " ", row.names(outcomes7), fixed=T)

names(outcomes7) <- gsub("_", " ", names(D_doses)[immune.markers.log])
names(outcomes7) <- gsub(".", " ", names(outcomes7), fixed=T)

kable(outcomes7, row.names = TRUE, caption = "Compare Base to Base, Dose, Vacc - Binary Viral Load with Binned Dose Score", align = "c", escape = F) %>% kable_styling(latex_options=c(position = "HOLD_position")) %>% column_spec(1, width = "6em") %>% column_spec(c(2:(length(immune.markers.log)+1)), width = "3em")
```




